%This thesis investigated whether pre-classifying mobile network cells based on traffic behaviour can improve forecasting performance. A statistical classification method was developed using a t-test on holiday and non-holiday traffic differences, separating cells into seasonal and non-seasonal groups. Forecasting models including ARIMA/SARIMAX, OLS, and XGBoost were then evaluated on these groups. \\

%The classification results show clear structure across geographic site types. Urban cells are consistently classified as non-seasonal (52/52), while recreational and suburban cells show more mixed behaviour. Overall, 88 out of 91 cells in urban and recreational categories align with their expected classification based on geographic structure. Importantly, this separation was achieved purely from traffic behaviour, without using geographical location as an input, demonstrating that the method captures meaningful patterns in a data-driven and location-agnostic manner. \\

%It should be noted that this study used geographical location as a proxy for ground truth when evaluating the classification results. However, geographic category does not necessarily determine traffic behaviour, and the two may diverge in practice. A clear example is R3, which despite being geographically labeled as a recreational site, shows low holiday sensitivity similar to urban cells (Figure~\ref{fig:box_plot_holiday_avg_site}). Conversely, U3, despite being geographically labeled as urban, exhibits a daily activity pattern closer to those of recreational cells (Figure~\ref{fig:avg_hourly_per_site}), with low daytime activity and a pronounced evening peak. This suggests that that relying solely on location-based grouping may introduce misclassification. In the absence of true ground truth labels at the cell level, the agreement between the classification and geographic categories should therefore be interpreted as supporting evidence rather than definitive validation. \\

%In the forecasting task, all model-based approaches significantly outperform naive baselines, with error reductions of approximately 19\% compared to the best naive model. Best-performing models achieve RMSE values in the range of 1.45--1.70 across all datasets, compared to naive baseline RMSE values of 1.77--2.58. This confirms that the dataset contains strong temporal structure that can be learned using statistical and machine learning methods. \\

%Among all models, OLS and XGBoost achieve comparable performance overall, while ARIMA/SARIMAX is consistently the weakest. Paired statistical testing across cells confirms that ARIMA performs significantly worse than both OLS and XGBoost in every dataset split. The relationship between OLS and XGBoost depends on the subgroup: the two models are statistically indistinguishable on seasonal cells (RMSE of 1.45 and 1.46 respectively), while XGBoost holds a small but statistically significant advantage over OLS on non-seasonal cells (1.56 vs 1.60) and across all cells combined (1.53 vs 1.56). Although this advantage is statistically significant, the absolute difference remains small, and OLS achieves this performance using a simple, interpretable linear model rather than an ensemble of decision trees. This suggests that while a complex, non-linear model such as XGBoost can extract a modest additional benefit from pooling data across cells, a well engineered linear model captures nearly all of the same predictive signal while remaining considerably more explainable and easier to deploy and maintain in practice. Overall, this indicates that strong feature engineering, particularly lagged values and rolling statistics, captures most of the predictive signal regardless of model choice, and that the added complexity of a model like XGBoost is only weakly justified by the resulting gains in this setting. \\

%The impact of dataset segmentation is limited. Training separate models for seasonal and non-seasonal cells in the statistical models only results in marginal improvements compared to a global model, and only for the seasonal subset, which represents the smaller minority group. When a subgroup is already well represented in the global training data, as is the case for non-seasonal cells at 69\% of all samples, group-specific training provides little to no additional benefit. This suggests that pre-classification is most valuable when a meaningful class imbalance exists, which does not appear to be the case in this dataset. \\

%Overall, the results show that a simple pipeline combining statistical classification and feature-driven forecasting models, whether linear or tree-based, is sufficient to achieve strong performance in this setting when comparing to baseline models. Feature engineering plays a more important role than model complexity or segmentation of the data, although model choice still produces small, statistically significant differences depending on the seasonality of the subgroup. \\

This thesis tested a specific assumption underlying conventional network planning,  that geographic location, specifically whether a cell is urban or recreational, is a reliable proxy for its seasonal traffic behaviour. Rather than taking this assumption at face value, a data-driven classification method was applied that used only traffic data, with no knowledge of a cell's location, to test whether it would independently recover the same urban/recreational grouping. \\

It largely did. The t-test classification based on holiday traffic differences agreed with the geographic labels for 88 of 91 cells, with all urban cells correctly identified as non-seasonal and the large majority of recreational cells correctly identified as seasonal. The direction of the effect also matched expectations: recreational sites consistently showed higher traffic during holiday periods, consistent with increased leisure and travel activity, while urban sites consistently showed lower traffic during holidays. This supports the planning assumption that geography is a meaningful determinant of seasonality, while also showing that this property can be verified directly from traffic data rather than taken on faith. The few cells that broke from their expected category (R3, U3) indicate that geography is not a perfect predictor and that some individual cells behave more like the opposite category, information that would be missed by a purely location-based grouping. \\

The second question, whether grouping cells this way improves forecasting, returned a more limited answer. Forecasting models trained separately on seasonal and non-seasonal the subsets only marginally outperformed a single global model, with the benefit confined to the smaller seasonal subgroup. The larger driver of forecasting accuracy was feature engineering: once lagged values and rolling statistics were included, a simple linear model performed on par with a substantially more complex machine learning model. Segmentation by seasonality, validated the geographic assumption but was not the primary factor for improving forecasts in this dataset. \\

Future work could extend the classification approach beyond a simple seasonal versus non-seasonal split. For example, additional categories could be introduced for event and transportation driven cells, or cells could be scored on a continuous scale of seasonality rather than sorted into two fixed groups. Another direction could be classifying cells by the dominant device type or usage behaviour driving their traffic, such as distinguishing cells dominated by smartphone usage from those with a higher share of smart TV or streaming traffic, since these are likely to produce distinct volume and timing patterns. Unsupervised methods could be explored as an alternative to the t-test used in this thesis for identifying traffic regimes. Finally, deep learning models such as LSTMs or CNNs could be tested, as they may capture temporal and spatial patterns across cells more effectively than the linear and tree-based models used here. \\