

Mobile networks are under constant pressure to deliver reliable service, while the resources required to do so, including frequency spectrum, infrastructure, and capacity, remain costly. When operators fail to anticipate future traffic demand, the consequences are wasted capacity, degraded user experience, and inefficient infrastructure investment. Beyond immediate performance, accurate traffic forecasting enables operators to make informed decisions about where and when to upgrade network infrastructure, ensuring that capital is directed toward cells where demand genuinely warrants it \\

Mobile network cells do not follow a single traffic pattern. Geographic location strongly influences both the level and timing of traffic. These differences are driven by human behavior, such as commuting and recreational activity, as well as network effects like load shifting between nearby cells. As a result, forecasting across all cells is a diverse time-series problem. \\

Because of this variation, building a robust forecasting model is not straightforward. One approach is to model each cell independently. This preserves local structure but is computationally expensive and can lead to overfitting, especially for cells with limited historical data. Another approach is to train a single global model across all cells. This allows shared representations of traffic patterns and can capture differences between cell types. However, it reduces interpretability and statistical consistency, since the model combines data from sites with different underlying traffic patterns. A third approach is to pre-classify cells based on their dominant traffic behavior, grouping cells with similar temporal structures. This reduces diversity within each group, making traffic patterns more consistent across cells within the same category. In this setting, simpler statistical models such as ARIMA or OLS can remain competitive when applied to more consistent subsets of cells. \\

%Different types of cells exhibit distinct traffic patterns. cell close to event locations, such as stadiums and exhibition centers, could experience long periods of low activity interrupted by short, sharp peaks triggered by external events. These series are non-stationary, since their statistical properties change abruptly during event windows. Traditional models such as a linear regression might struggle to capture the non linear patterns of these spikes \\

%Seasonal or recreational cells, such as ski resorts, coastal regions, and holiday destinations, follow smoother and more regular seasonal cycles. These patterns are visible across months, or years, with recurring peaks during specific periods. Here the traffic is typically not stable week to week, but might show seasonal cycles so its stable year to year. \\

%Urban and commercial cells tend to exhibit a more stable structure. Traffic in these areas is driven by regular human activity patterns such as work commutes and daily routines. This produces strong and consistent daily cycles with clear autocorrelation at short lags such as 24 hours or weekly lags. These series are often closer to stationarity, making them well suited for simpler forecasting models, such as a linear regression, where we assume traffic follows a constant trend over time.. \\

This work investigates whether pre-classifying cells based on traffic behavior can improve forecasting performance compared to a single global approach. Using historical data from TDC NET's mobile network, a data-driven classification method is developed to separate cells into seasonal and non-seasonal categories, and several forecasting models are evaluated within each group. \\
%Given this diversity, a single global model may struggle to represent all traffic behaviors effectively. this work investigates whether pre-classifying cells based on traffic behavior can improve forecasting performance compared to a single global approach. \\
 
The remainder of this work first describes the dataset and preprocessing steps, then introduces a method for classifying cells based on seasonality and traffic behavior, and finally evaluates several forecasting approaches, including OLS, ARIMA, and XGBoost, in terms of forecasting performance. \\ 